Stage-Splitting a Modal Language

### 2014-07-24 paper/semantics.tex
work on 3.2

\TODO clean up the next few paragraphs; unclear how much of it should go in the
paper

@@

Intuitively, the solution is to say that \verb|next {1+2}| is a value \emph{at
stage \bbone}, even though once we proceed to stage-\bbtwo\ evaluation, it will
no longer be a value. Therefore, we evaluate the pair to
\begin{lstlisting}
#2 (next {1+2}, 7)
\end{lstlisting}
and get \verb|7| as the answer.

To complicate matters, consider a situation where we must substitute a stage
\bbone\ value into some other expression. In the stage-\bbtwo\ term
\begin{lstlisting} 
prev{
  let x = (next {1+2}, 3+4) in
  next{ prev{#1 x} * prev{#1 x} * hold{#2 x} }
}
\end{lstlisting}
our stage-\bbone\ evaluation will clearly yield a term which is not fully
reduced---since \verb|1+2| is needed to compute the end result, but we cannot
reduce it to \verb|3| until stage \bbtwo.

On the other hand, if we simply treat \verb|(next {1+2}, 7)| as a value, then we
will duplicate the \verb|1+2| computation in the term, and in fact we will
evaluate it twice, in each computation of \verb|#1 (next {1+2}, 7)|.

Our solution is to hoist these residuals, etc.

\vspace{2em}
\TODO old stuff below here. merge with above text:

Our goal is to evaluate away all of the stage-\bbone\ portions of this code, but there is more than one reasonable way to do this.
One option, taken by \cite{davies96}, is to treat \verb|next{1+2}| as a value for the purposes of stage-\bbone~evaluation, 
consequently duplicating it during substitution for {\tt x}.  The two $\prev$ constructs eliminate the $\next$ wrappers,
leaving the contained stage-\bbtwo\ code in place.
This produces:
\begin{lstlisting}
(1+2) * (1+2) * 7
\end{lstlisting}

We take a different approach.  
Instead of duplicating the contents of the first $\next$ expression, we bind them to some variable (here, $\mathtt{y}$) and
duplicate only a reference to that variable.  This produces:
\begin{lstlisting} 
let y = 1+2 in y * y * 7
\end{lstlisting}

>>>>

@@

Intuitively, the solution is to say that \verb|next {1+2}| is a value \emph{at
stage \bbone}, even though it requires additional evaluation at stage \bbtwo.
Therefore, we evaluate the pair to
\begin{lstlisting}
#2 (next {1+2}, 7)
\end{lstlisting}
and get \verb|7| as the answer.

To complicate matters, consider a situation where we must substitute such a
incompletely-evaluated expression. The following stage-\bbtwo\ term has type
$\rmint$:
\begin{lstlisting} 
prev{
  let x = (next {1+2}, 3+4) in
  next{ prev{#1 x} * prev{#1 x} * hold{#2 x} }
}
\end{lstlisting}
Stage-\bbone\ evaluation will not fully reduce this term, because the answer
depends on the value of \verb|1+2|, but we cannot reduce this to \verb|3| until
stage \bbtwo.

If we simply treat \verb|(next {1+2}, 7)| as a value during stage \bbone, then
we will substitute it for the three occurrences of \verb|x| in the body of the
\verb|let| expression. Thus, the result of stage \bbone\ computation is
\begin{lstlisting} 
prev{
  next{ prev{next {1+2}} * prev{next {1+2}} * 7 }
}
\end{lstlisting}
When we evaluate this at stage \bbtwo, we will compute \verb|1+2| twice. Of
course, this is not particularly expensive, but in general, the duplicated
computation could be anything, even a recursive call to the function!

To avoid duplicating computations, we take a different approach.  Instead of
duplicating the contents of the first $\next$ expression, we bind them to some
variable (here, $\mathtt{y}$) and duplicate only a reference to that variable.
This produces:
\begin{lstlisting} 
let y = 1+2 in y * y * 7
\end{lstlisting}
